Music adaptation: Analysis

### 2022-10-07 analysis/baseline.Rmd
update stats for ioi, kot
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Comparison between baseline and test performance'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

@@

# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# stats
if (!require("afex")) {install.packages("afex"); require("afex")}
# plot
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
```

@@

ioi_trial <- dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category)]

ioi <- ioi_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics, Category)]
setorder(ioi, "SubNr", "Articulation", "Dynamics", "Category")
ioi
```

@@

```{r ioi-all}
# group mean
ioi_all <- ioi[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation, Dynamics, Category)]
ioi_all
```

@@

```{r ioi-all-plot}
ggplot(ioi, aes(x = Articulation, y = Mean, color = Dynamics))+ 
  geom_boxplot(position = position_dodge2(preserve = "single")) +
  geom_point(binaxis = "y", stackdir = "center", position = position_dodge(0.75)) +
  theme_pubr()
#ggboxplot(ioi, "Articulation", "Mean", color = "Dynamics", add = "jitter", xlab = "Articulation", ylab = "IOIs (ms)", title = "IOI")
```

@@

@@

@@

kot <- kot_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics, Category, Subcomponent1)]
setorder(kot, "SubNr", "Articulation", "Dynamics", "Category")
kot
```

@@

```{r kot-all-plot, echo = FALSE}
ggboxplot(kot[Subcomponent1 == "Legato"], "Articulation", "Mean", color = "Dynamics", add = "jitter", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Legato)")

ggboxplot(kot[Subcomponent1 == "Staccato"], "Articulation", "Mean", color = "Dynamics", add = "jitter", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Staccato)")
```

@@

1) Legato

```{r}
aov_ez(
  data = kot[Subcomponent1 == "Legato"],
  id = "SubNr",
  dv = "Mean",
  within = c("Category")
)
```

2) Staccato

```{r}
aov_ez(
  data = kot[Subcomponent1 == "Staccato"],
  id = "SubNr",
  dv = "Mean",
  within = c("Category")
)
```

# Dynamics
## 1. Individual

```{r vel, echo = FALSE}
# for each individual
vel_trial <- dt_vel[Subcomponent2 != "NA", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category, Subcomponent2)]

@@

```{r vel-all-plot,  echo = FALSE}
ggboxplot(vel[Subcomponent2 == "Forte"], "Dynamics", "Mean", color = "Articulation", add = "jitter", xlab = "Dynamics", ylab = "Velocity", title = "KV (Forte)")

ggboxplot(vel[Subcomponent2 == "Piano"], "Dynamics", "Mean", color = "Articulation", add = "jitter", xlab = "Dynamics", ylab = "Velocity", title = "KV (Piano)")
```

@@

```{r vel-diff, echo = FALSE}
# for each individual
vel_diff_trial <- dt_vel_diff[Subcomponent2 == "FtoP" | Subcomponent2 == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category, Subcomponent2)]

@@

```{r vel-diff-all-plot,  echo = FALSE}
ggboxplot(vel_diff[Subcomponent2 == "FtoP"], "Dynamics", "Mean", color = "Articulation", add = "jitter", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV (FtoP)")

ggboxplot(vel_diff[Subcomponent2 == "PtoF"], "Dynamics", "Mean", color = "Articulation", add = "jitter", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV (PtoF)")
```

>>>>

---
title: 'Music adaptation: Analysis'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

@@

# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# stats
if (!require("afex")) {install.packages("afex"); require("afex")}
if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
# plot
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
```

@@

ioi_trial <- dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI), CV = sd(IOI)/mean(IOI)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category)]

ioi <- ioi_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), CV = mean(CV)), by = .(SubNr, Articulation, Dynamics, Category)]
setorder(ioi, "SubNr", "Articulation", "Dynamics", "Category")
ioi
```

@@

```{r ioi-all}
# group mean
ioi_all <- ioi[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean), CV = mean(CV)), by = .(Articulation, Dynamics, Category)]
ioi_all
```

@@

```{r ioi-all-plot}
ggplot(ioi, aes(x = Articulation, y = Mean, color = Dynamics))+ 
  geom_boxplot(position = position_dodge2(preserve = "single")) +
  geom_point(position = position_dodge(0.75)) +
  theme_pubr()
#ggboxplot(ioi, "Articulation", "Mean", color = "Dynamics", add = "jitter", xlab = "Articulation", ylab = "IOIs (ms)", title = "IOI")
```

@@

1) Articulation x Dynamics

@@

2) Category: All

```{r}
aov_ez(
  data = ioi,
  id = "SubNr",
  dv = "Mean",
  within = c("Category")
)
```

3) Category: both vs. none vs. baseline

```{r}
aov_ez(
  data = ioi[Category == "both" | Category == "none" | Category == "baseline"],
  id = "SubNr",
  dv = "Mean",
  within = c("Category")
)
```

4) Variability

```{r}
aov_ez(
  data = ioi,
  id = "SubNr",
  dv = "SD",
  within = c("Category")
)
```

@@

kot <- kot_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics, Category, Subcomponent1)]
setorder(kot, "SubNr", "Subcomponent1", "Articulation", "Dynamics", "Category")
kot
```

@@

```{r kot-all-plot, echo = FALSE}
ggplot(kot[Subcomponent1 == "Legato"], aes(x = Articulation, y = Mean, color = Dynamics))+ 
  geom_boxplot(position = position_dodge2(preserve = "single")) +
  #geom_point(position = position_dodge(0.75)) +
  facet_grid( ~ Subcomponent1) +
  theme_pubr()

ggplot(kot[Subcomponent1 == "Staccato"], aes(x = Articulation, y = Mean, color = Dynamics))+ 
  geom_boxplot(position = position_dodge2(preserve = "single")) +
  #geom_point(position = position_dodge(0.75)) +
  facet_grid( ~ Subcomponent1) +
  theme_pubr()

#ggboxplot(kot[Subcomponent1 == "Legato"], "Articulation", "Mean", color = "Dynamics", add = "jitter", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Legato)")

#ggboxplot(kot[Subcomponent1 == "Staccato"], "Articulation", "Mean", color = "Dynamics", add = "jitter", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Staccato)")
```

@@

## 4. Stats
### Legato

1) Articulation x Dynamics

```{r}
aov_ez(
  data = kot[Subcomponent1 == "Legato" & Category != "baseline"],
  id = "SubNr",
  dv = "Mean",
  within = c("Articulation", "Dynamics")
)
```
2) Category: All

```{r}
leg_ca_all_aov <- aov_ez(
  data = kot[Subcomponent1 == "Legato"],
  id = "SubNr",
  dv = "Mean",
  within = c("Category")
)
leg_ca_all_aov

pairs(emmeans(leg_ca_all_aov, "Category"), adjust = "tukey")
```

3) Category: both vs. dyn_only vs. baseline

```{r}
leg_ca_par_aov <- aov_ez(
  data = kot[Subcomponent1 == "Legato" & Category == "both" | Subcomponent1 == "Legato" & Category == "dyn_only" | Subcomponent1 == "Legato" & Category == "baseline"],
  id = "SubNr",
  dv = "Mean",
  within = c("Category")
)
leg_ca_par_aov

pairs(emmeans(leg_ca_par_aov, "Category"), adjust = "tukey")
```

### Staccato

1) Articulation x Dynamics

```{r}
aov_ez(
  data = kot[Subcomponent1 == "Staccato" & Category != "baseline"],
  id = "SubNr",
  dv = "Mean",
  within = c("Articulation", "Dynamics")
)
```
2) Category: All

```{r}
sta_ca_all_aov <- aov_ez(
  data = kot[Subcomponent1 == "Staccato"],
  id = "SubNr",
  dv = "Mean",
  within = c("Category")
)
sta_ca_all_aov

pairs(emmeans(sta_ca_all_aov, "Category"), adjust = "tukey")
```

3) Category: both vs. dyn_only vs. baseline

```{r}
sta_ca_par_aov <- aov_ez(
  data = kot[Subcomponent1 == "Staccato" & Category == "both" | Subcomponent1 == "Staccato" & Category == "dyn_only" | Subcomponent1 == "Staccato" & Category == "baseline"],
  id = "SubNr",
  dv = "Mean",
  within = c("Category")
)
sta_ca_par_aov

pairs(emmeans(sta_ca_par_aov, "Category"), adjust = "tukey")
```
# Dynamics
## 1. Individual

```{r vel, echo = FALSE}
# for each individual
dt_vel$Articulation <- factor(dt_vel$Articulation, c("Present", "Absent", "Baseline"))
dt_vel$Dynamics <- factor(dt_vel$Dynamics, c("Present", "Absent", "Baseline"))

vel_trial <- dt_vel[Subcomponent2 != "NA", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category, Subcomponent2)]

@@

```{r vel-all-plot,  echo = FALSE}
ggplot(vel, aes(x = Dynamics, y = Mean, color = Articulation))+ 
  geom_boxplot(position = position_dodge2(preserve = "single")) +
  geom_point(binaxis = "y", stackdir = "center", position = position_dodge(0.75)) +
  facet_grid( ~ Subcomponent2) +
  theme_pubr()

#ggboxplot(vel[Subcomponent2 == "Forte"], "Dynamics", "Mean", color = "Articulation", add = "jitter", xlab = "Dynamics", ylab = "Velocity", title = "KV (Forte)")

#ggboxplot(vel[Subcomponent2 == "Piano"], "Dynamics", "Mean", color = "Articulation", add = "jitter", xlab = "Dynamics", ylab = "Velocity", title = "KV (Piano)")
```

@@

```{r vel-diff, echo = FALSE}
# for each individual
dt_vel_diff$Articulation <- factor(dt_vel_diff$Articulation, c("Present", "Absent", "Baseline"))
dt_vel_diff$Dynamics <- factor(dt_vel_diff$Dynamics, c("Present", "Absent", "Baseline"))

vel_diff_trial <- dt_vel_diff[Subcomponent2 == "FtoP" | Subcomponent2 == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category, Subcomponent2)]

@@

```{r vel-diff-all-plot,  echo = FALSE}
ggplot(vel_diff, aes(x = Dynamics, y = Mean, color = Articulation))+ 
  geom_boxplot(position = position_dodge2(preserve = "single")) +
  geom_point(binaxis = "y", stackdir = "center", position = position_dodge(0.75)) +
  facet_grid( ~ Subcomponent2) +
  theme_pubr()

#ggboxplot(vel_diff[Subcomponent2 == "FtoP"], "Dynamics", "Mean", color = "Articulation", add = "jitter", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV (FtoP)")

#ggboxplot(vel_diff[Subcomponent2 == "PtoF"], "Dynamics", "Mean", color = "Articulation", add = "jitter", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV (PtoF)")
```
